---
title: "EDA On Titanic Dataset"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

Exploratory data analysis on the titanic dataset. 

```{r}
library(ggplot2)
train <- read.csv("C:/Users/hp/Downloads/train.csv")
titanic<- train
Pclass<-as.factor(titanic$Pclass)
Sex<- as.factor(titanic$Sex)
titanic$Survived<- as.factor(titanic$Survived)
```

We can see the summary statistics of the titanic dataset.

```{r}
summary(titanic)
str(titanic)

```
Survival rate in the titanic:
```{r}
ggplot(data= titanic, aes(x= Survived))+ geom_bar() + theme_linedraw()+ 
  labs(y= "Count of passengers ", title = "Survival rate of Passengers")
```

We see that unfortunately most of the people died in the tragedy. We can see the exact 
percentage as well.
```{r}
prop.table(table(titanic$Survived))
```

61.1% of the people did not survive in the titanic
We can also see some other insights from the given dataset. 
```{r}
table(titanic$Sex)
prop.table(table(titanic$Sex))
ggplot(data= titanic, aes(x= Survived, fill= Sex))+ geom_bar() + theme_linedraw() +
  labs(y="Count of Passengers", title = "Survival rate of Passengers by Sex")

```

We see from the above visualization that unfortunately, most of the males did not survive in the ship. This means the females were given more priority than the males.
```{r}
ggplot(data= titanic, aes(x= Age))+ geom_histogram(binwidth = 5)+ theme_linedraw()+
  labs(y= "Count of passengers", title = "Age distribution of passengers")

ggplot(data= titanic, aes(x= Age))+ geom_density()+ theme_linedraw()+
  labs(y= "Count of passengers", title = "Age distribution of passengers")


ggplot(data= titanic, aes(x= Age, fill= Survived)) +geom_density(alpha= 0.5)
```

We see that the children survived more than the people of greater age. Unfortunately, people
of the age 70-80 years of age did not survive much.
We can also see what the Passenger class has to do with the survival rate of the passengers.
```{r}
ggplot(data= titanic, aes(x= Pclass))+ geom_bar() + theme_linedraw()+ 
  labs(x= "Passenger Class", y= "Count of Passengers" )
```

We see that most of the passengers are from the passenger class 3.
```{r}
ggplot(data= titanic, aes(x= Pclass, fill= Survived))+ geom_bar() + theme_linedraw()+ 
  labs(x= "Passenger Class", y= "Count of Passengers", title = "Passenger survival by class" )
```

Here, we see that more percentage of people survived in the passenger class 1 survived than
passenger class 2 and passenger class 3.Survival Percentage of people in class 2 is more
than the survival percentage in class 3. One possibility is that the life boats could be 
closer to class 1 than class 2 and class 3. Also, life boats could be closer in class 2 
relative to class 3.

We can also see the survival rate of the passengers by class and by the sex.
```{r}
ggplot(data= titanic, aes(x= Pclass, fill= Survived)) +theme_linedraw() + geom_bar() + 
  facet_wrap(Sex)
```

We see that almost all the females in class 1 and class 2 survived more than the females in
class 3. On the other hand, since males were given less preference than females, the 
survival rate of males is subtantially lower than the females. Also, here we see that due
to the biasness by class, the males who were in the passenger class 3 suffered the most.

```{r}
ggplot(data= titanic, aes(x= Age, fill= Survived))+ theme_linedraw() + geom_histogram()+ 
  facet_wrap(Pclass~ Sex)

ggplot(data= titanic, aes(x= Age, fill= Survived))+ theme_linedraw() + 
  geom_density(alpha= 0.5)+ facet_wrap(Pclass~ Sex)
```

The above two plots give a lot of information about the data. The age factor is also 
introduced.

```{r}
ggplot(data= titanic, aes(x= Embarked, fill= Survived)) + geom_bar() + theme_linedraw() +
  labs(y= "count of passengers", title = "Survival rate ")
```

We see that the point of embarkment of most of the passengers was "S" and more than 50% of
the people who embarked from "S" did not survive.

