---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

We will do some analysis on the gapminder dataset.
```{r}
library(dplyr)
library(ggplot2)
library(gapminder)
str(gapminder)
summary(gapminder)

```

We will do some comparisons between India and Singapore.
```{r}
df2<- gapminder %>% filter(country == "India" | country == "Singapore")
ggplot(data= df2, aes(x= gdpPercap, y= lifeExp, col = country))+ geom_line()+
  geom_point()

ggplot(data= df2, aes(x= lifeExp, y= year, col= "red"))+  geom_point() +
  facet_wrap(df2$country)
ggplot(data= df2, aes(x= gdpPercap, y= year, col= "red"))+  geom_point()+ geom_line()+
  facet_wrap(df2$country)

ggplot(data= df2, aes(x= pop, y= year, col= "red"))+  geom_point()+ geom_line() +
  facet_wrap(df2$country)
```


We have compared the life expectancy, per capita GDP and population of India and Singapore. From the above visualizations, we see that Singapore is lot more developed than India.

```{r}
df3<- gapminder %>% filter(year == 2002)
```

We have made a data frame which contains all the data from the year 2002 only.

```{r}
ggplot(data= df3, aes(x= lifeExp)) + geom_histogram(binwidth = 5) + 
  labs(x=" Life Expectancy", y= "Count of Countries ") + theme_linedraw()
```

We see that most of the countries have life expectancy between 65 to 85 years.

We can also see the visualization of life expectancy with respect to the continent.
```{r}
ggplot(data= df3, aes(x= lifeExp)) + geom_histogram(binwidth = 5) + facet_wrap(df3$continent)+
  theme_linedraw() + labs(x= "Life Expectancy", y= "Count of Countries")
```

We see that Africa does not have much life expectancy relative to the other continents
Most of the countries in Africa have life expectancy between 45 - 55 years. 
Americas and Euprope have high life expectancy rates.

```{r}
q<- ifelse(df3$lifeExp>50, "yes", "no")
df3$life50<- q
str(df3)
summary(df3)
ggplot(data= df3, aes(x= continent, fill= life50))+  geom_bar() + theme_linedraw() +
  labs(y= "Count of Countries", title = "life Expectancy>50 or not")
```

We see that more than 20 countries in Africa have less than or equal to 50 years of life
expectancy.

```{r}
summary(df3$gdpPercap)
w<- ifelse(df3$gdpPercap>5000, "yes", "no")
df3$gdp5000<-w
```

We have made a new variable in which the value is "yes" if GDP per capita is greater than 
5000 and "no" if GDP per capita is less than 5000.

```{r}
ggplot(data = df3, aes(continent, fill= gdp5000))+ geom_bar() +theme_linedraw() +
  labs(y= "Count of Countries", title = "GDP>5000 or not")
```

We see that most of the countries in Africa have GDP per capita less than 5000. 
Gross domestic product (GDP) is a monetary measure of the market value 
of all the final goods and services produced in a period of time, often annually.
GDP per capita is the GDP of the country divided by its population.

We can also find the GDP of each country by multiplying population of each country 
with the per capita GDP.

```{r}
e<- df3$pop * df3$gdpPercap
df3$gdp<- e
str(df3)
summary(df3)

```

